\documentclass{article}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[margin=1in]{geometry}
\usepackage{newtxmath,newtxtext}

\newcommand{\HRule}{\rule{\linewidth}{0.5mm}}
\newcommand{\Hrule}{\rule{\linewidth}{0.3mm}}

\makeatletter% since there's an at-sign (@) in the command name
\renewcommand{\@maketitle}{%
  \parindent=0pt% don't indent paragraphs in the title block
  \centering
  {\Large \bfseries\textsc{\@title}}
  \HRule\par%
  \textit{\@author \hfill \@date}
  \par
}
\makeatother% resets the meaning of the at-sign (@)

\title{Statement of Purpose}
\author{Shreyas Kalvankar}
\date{M.S. Applicant}

\begin{document}
  \maketitle% prints the title block
  \thispagestyle{empty}

  \vspace{5pt}
\hspace{0.25in}My research interests lie in uncovering the mathematics that drives Machine
  Learning and deep learning and their theoretical nature. What are the
  theoretical foundations explaining the success of deep networks? Are there
  inherent limits to their expressiveness? What are the potential applications
  of learned representations in various domains, such as computer vision,
  natural language processing, and speech recognition? How do we address the
  challenges of interpretability and explainability in these learned
  representations? I aim to explore these problems to develop an understanding
  of the theoretical limits of Machine Learning. I am particularly interested in
  understanding and applying these algorithms across engineering and natural
  sciences to build intelligent systems and conduct research in physics and
  mathematics to solve problems such as approximating numerical computations
  using neural networks, astronomical simulations, assisted theorem proving,
  etc. My unwavering passion for engineering and research is evident in my
  academic achievements during my undergraduate years, reflected in my
  transcripts and my active involvement in extracurricular activities. My
  journey has underscored the importance of further specialization and
  enrichment, which I believe a master's program can provide, laying the
  foundation for a future Ph.D. pursuit. In this regard, I am convinced that
  North Carolina State University is the ideal place to nurture my potential.

  \vspace{5pt}
\hspace{0.25in}During my bachelor’s I was extremely fascinated by astrophysics and its
intersection with computer science. In my junior year, I led a team of three and
delved into the research on neural network applications in galaxy morphology
classification. Identifying galaxy morphologies has important implications in
many astronomical tasks, e.g., studying galaxy evolution. The recent data influx
in astronomy necessitates a robust and automated system for processing large
amounts of images. We aimed to set up a 7-class classification system that
classified galaxy images using CNNs, which can surpass existing benchmarks. I
was responsible for establishing a robust training pipeline and optimizing
various neural network architectures from the ground up. Our collective efforts
outperformed the second-best submission on the Kaggle Public Leaderboard. I
gained a profound understanding of the intricacies of applying deep learning to
various tasks and explored aspects such as hyperparameter tuning, debugging,
troubleshooting, and custom model design—each underscored by a need for a
profound grasp of the underlying theory. My role as the project lead was a
significant learning experience that taught me various aspects of management,
communication, and project planning. Through this project, I honed my research
skills, nurtured my critical thinking abilities, and sharpened my
decision-making capabilities. 

  \vspace{5pt}
\hspace{0.25in}For my bachelor’s thesis, I worked on utilizing Conditional GANs for
astronomical image colorization. Space archives are filled with large amounts of
low-quality, greyscale images, many of which go unnoticed. I was interested in
utilizing generative models for creating aesthetically pleasing representations
of celestial scenes. Generative models are attractive because they enable us to
better understand, create, and work with data, having many practical and
theoretical implications. In my work, I focused on tasks like image-to-image
translation and style transfer, where I realized the need for distributional
robustness in my application to maintain quality in my output images, even in
varying conditions. GANs often deal with high-dimensional data and are used for
various conditional image generation tasks, hence, it is important to study
their optimization landscape and convergence guarantees. In my ongoing academic
journey, I’m equally interested in gaining a theoretical understanding of how
various neural network models generalize, especially in complex,
high-dimensional spaces when provided with different conditions or inputs.
Undergraduate studies seldom comprehensively cover advanced machine learning
theory. It is necessary to understand the generalization bounds of different
learning algorithms to prepare for a successful research career in ML. I wish to
study these topics during my graduate studies at NCSU as a part of the Special
Topics in Computer Science course (CSC-295 \& CSC-495) and further my understanding.

  \vspace{5pt}
\hspace{0.25in}The Galaxy Morphology Classification project proved to be a stepping stone in
connecting astrophysics to computer science. Soon after that project, I began
contributing to EinsteinPy, an open-source Python package designed to address
issues in General Relativity and gravitational physics. My work specifically
involved incorporating various symbolic computations, such as the
Reissner–Nordström metric and calculations for the event horizon and ergosphere
of a Kerr-Newmann black hole. While symbolic computations are valuable in
controlled environments, real-world problems often necessitate approximate
rather than exact solutions. For this reason, the potential for deep learning in
simulating such complex system behaviors is vast, e.g., in surrogate modeling
where we can use deep learning to approximate behaviors in complex systems
allowing for faster evaluations when original simulation using numerical methods
is computationally expensive. In simulations, gravitational interactions are
currently computed using hierarchical tree algorithms like the Barnes-Hut \& Fast
Multipole Method. I am interested in leveraging AI and Machine Learning for this
task in the long term. One possible approach is to model the gravitational
interactions of various particles as nodes of a graph and learn the patterns as
they evolve over time by using Graph Neural Networks. I believe Prof. Xiaorui
Liu’s expertise in Graph Neural Networks can help me immensely in my future
research pursuits. Owing to an overlap of interests in large-scale optimization,
ML with graphs, and robust \& certifiable ML, he would be a compelling choice as a master’s advisor.

  \vspace{5pt}
\hspace{0.25in}After completing my degree, I embarked on my professional journey, starting as a
Machine Learning Engineer and later transitioning to a Machine Learning
Consultant role at the startup Relfor Labs. My primary focus revolved around
applying deep learning to audio data analysis and classification by devising
novel architectures. I worked on designing a system for pre-processing large
amounts of raw audio data by converting it into mel spectrograms, refining the
labels. I set up an end-to-end training pipeline to streamline experimentation.
Throughout my work on this and various other ML and data-intensive projects, I
have been interested in setting up self-sustaining models which can keep
learning in the long term. However, I have faced several challenges in my
attempts to achieve this which include data quality and consistency issues,
model drift, and version control. My experience in developing ML models and
designing ML pipelines, along with the challenges I faced make the prospect of
working with Prof. Jung-Eun Kim to build sustainable and autonomous systems
extremely fascinating.

  \vspace{5pt}
\hspace{0.25in}The scientific landscape has been consistently evolving, but many longstanding
problems have eluded complete solutions for many years. Looking into the future,
I wish to make a difference in this significant era. My determination to pursue
a research career is unwavering, and I aspire to work in an academic setting and
eventually pursue a Ph.D. The vibrant community of students and researchers at
NCSU is the perfect place for me to nurture myself and further my commitment to
help expedite this scientific advancement. Thank you for considering me as a
prospective student at your university.

\end{document}
