\documentclass{article}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[left=1in, right=1in, top=1in, bottom=1in]{geometry}
\usepackage{newtxmath,newtxtext}

\newcommand{\HRule}{\rule{\linewidth}{0.5mm}}
\newcommand{\Hrule}{\rule{\linewidth}{0.3mm}}

\makeatletter% since there's an at-sign (@) in the command name
\renewcommand{\@maketitle}{%
  \parindent=0pt% don't indent paragraphs in the title block
  \centering
  {\Large \bfseries\textsc{\@title}}
  \HRule\par%
  \textit{\@author \hfill \@date}
  \par
}
\makeatother% resets the meaning of the at-sign (@)

\title{Statement of Purpose}
\author{Shreyas Kalvankar}
\date{Ph.D. Applicant}

\begin{document}
  \maketitle% prints the title block
  \thispagestyle{empty}
\vspace{5pt}
\hspace{0.25in}My research interests lie in uncovering the mathematics that
drives Machine Learning and deep learning and their theoretical nature. What are
the theoretical foundations explaining the success of deep networks? Are there
inherent limits to their expressiveness? What are the potential applications of
learned representations in various domains, such as computer vision, natural
language processing, and speech recognition? How do we address the challenges of
interpretability and explainability in these learned representations? I aim to
explore these problems to develop an understanding of the theoretical limits of
Machine Learning. I am particularly interested in understanding and applying
these algorithms across engineering and natural sciences to build intelligent
systems and conduct research in physics and mathematics to solve problems such
as approximating numerical computations using neural networks, astronomical
simulations, assisted theorem proving, etc. My unwavering passion for
engineering and research is evident in my academic achievements during my
undergraduate years, reflected in my transcripts and my active involvement in
extracurricular activities. My journey has underscored the importance of further
specialization and enrichment required for a successful research career. In this
regard, I am convinced that Caltech is the ideal place to nurture my
potential.\vspace{5pt}

\hspace{0.25in}During my bachelor's I was extremely fascinated by astrophysics
and its intersection with computer science. In my junior year, I led a team of
three and delved into the research on neural network applications in galaxy
morphology classification. Identifying galaxy morphologies has important
implications in many astronomical tasks, e.g., studying galaxy evolution. The
recent data influx in astronomy necessitates a robust and automated system for
processing large amounts of images. We aimed to set up a 7-class classification
system that classified galaxy images using CNNs, which can surpass existing
benchmarks. I was responsible for establishing a robust training pipeline and
optimizing various neural network architectures from the ground up. Our
collective efforts outperformed the second-best submission on the Kaggle Public
Leaderboard. I gained a profound understanding of the intricacies of applying
deep learning to various tasks and explored aspects such as hyperparameter
tuning and custom model design for converging to global optimum—each underscored
by a need for a profound grasp of the underlying theory. Training large-scale
models requires enormous compute, but has been made efficient owing to their
parallel nature. Dr. Anima Anandkumar's work in the analysis and design of
tensor algorithms is extremely relevant to my interests in developing systems
for physical sciences which often deal with high-dimensional data. I am
particularly interested to explore and design efficient ways to escape local
minimums, a problem generally faced during training complex models. I would like
to work with her to gain an in-depth understanding of the subject.\vspace{5pt}

\hspace{0.25in}The Galaxy Morphology Classification project proved to be a
stepping stone in connecting astrophysics to computer science. Soon after that
project, I began contributing to EinsteinPy, an open-source Python package
designed to address issues in General Relativity and gravitational physics. My
work specifically involved incorporating various symbolic computations, such as
the Reissner–Nordström metric and calculations for the event horizon and
ergosphere of a Kerr-Newmann black hole. While symbolic computations are
valuable in controlled environments, real-world problems often necessitate
approximate rather than exact solutions. This led me to investigate the
processes used for various physical simulations where numerical methods are
widely popular. The potential for deep learning in simulating complex system
behaviors is vast, e.g., in surrogate modeling where we can use deep learning to
approximate behaviors in complex systems allowing for faster evaluations when
original simulation using numerical methods is computationally expensive. In
simulations, gravitational interactions are currently computed using
hierarchical tree algorithms like the Barnes-Hut \& Fast Multipole Method. I am
interested in leveraging AI and Machine Learning for this task in the long term.
One possible approach is to model the gravitational interactions of various
particles as nodes of a graph and learn the patterns as they evolve over time by
using Graph Neural Networks. My interests in these topics are multi-faceted and
at the center of multiple fields like astronomy, gravitational physics \& machine
learning. Dr. Katherine Bouman's research in computational imaging is extremely
inspiring. I believe our research interests overlap greatly and my background in
applying machine learning to physics provides me an opportunity to contribute to
her ongoing research. Another interesting problem is the simulation of
hydrodynamic processes involved in galaxy formation using deep learning. Dr.
Oscar Bruno and Dr. Elizabeth Carlson's individual expertise in high-performance
numerical PDE solvers and high-performance computing would guide me tremendously
while exploring this problem in further detail. I would be honored to have an
opportunity to work with them.\vspace{5pt}

\hspace{0.25in}For my bachelor's thesis, I worked on utilizing Conditional GANs
for astronomical image colorization. Space archives are filled with large
amounts of low-quality, greyscale images, many of which go unnoticed. I was
interested in utilizing generative models for creating aesthetically pleasing
representations of celestial scenes. Generative models are attractive because
they enable us to better understand, create, and work with data, having many
practical and theoretical implications. In my work, I focused on tasks like
image-to-image translation and style transfer, where I realized the need for
distributional robustness in my application to maintain quality in my output
images, even in varying conditions. GANs often deal with high-dimensional data
and are used for various conditional image generation tasks, hence, it is
important to study their optimization landscape and convergence guarantees. In
my ongoing academic journey, I'm equally interested in gaining a theoretical
understanding of how various neural network models generalize, especially in
complex, high-dimensional spaces when provided with different conditions or
inputs. In this regard, Dr. Yang Song's expertise in generative models makes him
interesting to me as an advisor. Dr. Oscar Leong's insights in inverse problems
and machine learning complement my interests perfectly, and I would like to work
with him on these problems.  \vspace{5pt}

\hspace{0.25in}For the past two years, I have been working as an ML consultant
at Relfor Labs, an AI startup. My professional experience in Machine Learning
has further motivated me to pursue graduate studies to kick-start a research
career. The scientific landscape has been consistently evolving, but many
longstanding problems have eluded complete solutions for many years. Looking
into the future, I wish to make a difference in this significant era. My
determination to pursue a research career is unwavering, and I aspire to
eventually work in an academic setting as a professor. The vibrant community of
students and researchers at Caltech is the perfect place for me to nurture
myself and further my commitment to help expedite this scientific advancement.
Thank you for considering me as a prospective student at your university.

\end{document}
